Stability and Chaos in Multi-Agent Reinforcement Learning

### 2020-10-08 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
minor mods

@@

To verify experimentally the theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q-Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The aim is to determine the regions in which games learnt using Q-Learning converge to an equilibrium. The results of these experiments are shown in Figure \ref{fig:NumericalExperiments}, with parameters chosen to match those in the analytic assessment (Figure \ref{fig:theory}).
% By examination it is found that, in many cases the system oscillates about some equilibrium point in a low dimensional chaotic manner (i.e. the manner in which the action probabilities change is indistinguishable from random motion) \textbf{get some graphs for this}

@@

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure. Note in the following that for any heatmap in which $\tau$ is swept over, the roles of $\alpha$ and $\tau$ are swapped.
\begin{enumerate}
   \item Fix the parameters $\tau, \gamma$. The former is held at 0.05 and the latter at 0.1. For the heatmaps in which $\tau$ is swept over $\alpha$ is held fixed at 0.05.
   \item Initialise values of $\Gamma, \alpha$ (or $\tau$ appropriately). These will be swept over in the experiment.
\item Generate 15 payoff matrices for both agents by sampling from a multi-variate Gaussian 
(variables are the payoff elements) with mean zero and covariance parameterised by $\Gamma$.
\item For each of these payoff matrices, initialise a set of agents with random initial conditions (i.e., random action probabilities).
\item Allow both sets of agents to learn over a maximum of $1.5 \times 10^4$ iterations.
\item Keep track of the action probabilities over a window of 500 iterations. At the end of each window, determine the percentage difference between the maximum and minimum values of each strategy component
\item If the difference is less than 1 \% consider the game converged. Otherwise continue to the next window.
\item If the game reaches $1.5 \times 10^4$ iterations without satisfying the relative distance criterion, consider it to be non-convergent. Determine the fraction of these 15 games which have converged.

@@

We see in Figure \ref{fig:NumericalExperiments} that the stability of the system is highly dependent on
the value of $\alpha$ and $\tau$ but not on $\Gamma$.

@@

\begin{figure}[t]
    \includegraphics[width = 1.1 \linewidth, center]{Figures/a5e2var.png}
    \caption{Figures showing the trajectory of action probabilities for 2 players with 5 actions trained on a game of $\Gamma = -0.5$ and $\tau = 0.05$. (Left) $\alpha = 0.01$, (Right) $\alpha = 0.05$.}
    \label{fig:AlphaVariation}
\end{figure}

In Figure \ref{fig:AlphaVariation}, we explicitly plot the trajectories taken by the action probabilities as two agents are trained on a game of $\Gamma = -0.5$. In the left figure, we choose $\alpha = 0.01$ whilst in the right we choose $\alpha = 0.05$. This gives a qualitative assessment of the behaviour at both extremes parameters considered. We note that the typical behaviour of the system is that the system first moves towards a fixed point and then vary in a pseudo-random manner within a vicinity of the fixed point. Critically though, the amount of variation in the case $\alpha = 0.05$ is greater than that of $\alpha = 0.01$, and so it does not settle to a fixed point (i.e. is considered unstable).

@@

Finally, Figure \ref{fig:pVarianceExpt} plot the degree variation about the fixed point displayed as the number of players $p$ increases. The variation is determined by first allowing the game to iterate for 5000 steps, so that it can be assumed to have reached the vicinity of the fixed point. Then, the action probabilities are recorded for a further 5000 iterations. At the end of this second period, the variance of the actions are calculated as

\begin{equation*}
    V(t) = \frac{1}{N} \sum_i \frac{1}{5000} \sum_t x_i(t)^2 - \left[\frac{1}{5000} \sum_t x_i(t) \right]^2 .
\end{equation*}

@@

Due to computational restraints, which we discuss further in the discussion, we are only able to increase $p$ to a maximum value of 6 and must restrict also to $N = 2$. We maintain $\alpha = 0.02$ and $\tau = 0.05$ to match with the parameters used for Figure \ref{fig:pvariation}. Finally, in Figure \ref{fig:pVarianceExpt}, we are averaging over 10 choices of $\Gamma$ in the range $[-1, 0]$.

>>>>

\fb{it would be good to have some introduction to the section here.}

@@

To verify experimentally the theoretical results and to examine the
underlying structure of stability and chaos in multi-agent Q-Learning,
we perform a series of numerical experiments by varying the parameters
$\Gamma$ and $\alpha$ whilst keeping $\tau$ fixed. The aim is to
determine the regions in which games learnt using Q-Learning converge
to an equilibrium. The results of these experiments are shown in
Figure \ref{fig:NumericalExperiments}, with parameters chosen to match
those in the analytic assessment (Figure \ref{fig:theory}).
% By examination it is found that, in many cases the system oscillates about some equilibrium point in a low dimensional chaotic manner (i.e. the manner in which the action probabilities change is indistinguishable from random motion) \textbf{get some graphs for this}

@@

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure. Note in the following that for any heatmap in which $\tau$ is swept over, the roles of $\alpha$ and $\tau$ are swapped.
\begin{enumerate}
   \item Fix the parameters $\tau, \gamma$. The former is held at 0.05
     and the latter at 0.1. For the heatmaps in which $\tau$ is swept
     over, $\alpha$ is held fixed at 0.05.

   \item Initialise values of $\Gamma, \alpha$ (or $\tau$
     appropriately). These will be swept over in the experiment.

   \item Generate 15 payoff matrices for both agents by sampling from
     a multi-variate Gaussian (variables are the payoff elements) with
     mean zero and covariance parameterised by $\Gamma$.

   \item For each of these payoff matrices, initialise a set of agents
     with random initial conditions (i.e., random action
     probabilities).

   \item Allow both sets of agents to learn over a maximum of $1.5
     \times 10^4$ iterations.

   \item Keep track of the action probabilities over a window of 500
     iterations. At the end of each window, determine the percentage
     difference between the maximum and minimum values of each
     strategy component

   \item If the difference is less than 1 \% consider the game
     converged. Otherwise continue to the next window.

   \item If the game reaches $1.5 \times 10^4$ iterations without
     satisfying the relative distance criterion, consider it to be
     non-convergent.

     \item Determine the fraction of these 15 games which
     have converged.

@@

We see in Figure \ref{fig:NumericalExperiments} that the stability of the system is highly dependent on the values of $\alpha$ and $\tau$, but not on $\Gamma$.

@@

\begin{figure}[t]
    \includegraphics[width = 1.1 \linewidth, center]{Figures/a5e2var.png}
    \caption{Figures showing the trajectory of action probabilities for 2 players with 5 actions trained on a game with $\Gamma = -0.5$ and $\tau = 0.05$. (Left) $\alpha = 0.01$, (Right) $\alpha = 0.05$.}
    \label{fig:AlphaVariation}
\end{figure}

In Figure \ref{fig:AlphaVariation}, we explicitly plot the
trajectories taken by the action probabilities as two agents are
trained on a game with $\Gamma = -0.5$. In the left figure, we choose
$\alpha = 0.01$, whilst on the right we choose $\alpha = 0.05$. This
gives a qualitative assessment of the behaviour at both extremes
considered. We note that the typical behaviour of the
system is to first move towards a fixed point and then
vary in a pseudo-random manner within a vicinity of the fixed
point. Critically though, the amount of variation in the case for $\alpha
= 0.05$ is greater than that of $\alpha = 0.01$, and so it does not
settle to a fixed point (i.e., is considered unstable).

\fb{how can we say that in one case it converges to a fixed point and in the other it doesn't?}

@@

Finally, Figure \ref{fig:pVarianceExpt} plot the degree variation about the fixed point displayed as the number of players $p$ increases. The variation is determined by first allowing the game to iterate for 5000 steps, so that it can be assumed to have reached the vicinity of the fixed point. Then, the action probabilities are recorded for a further 5000 iterations. At the end of this second period, the variance of the actions are calculated as
%
\begin{eqnarray*}
    V(t) & = & \frac{1}{N} \sum_i \frac{1}{5000} \sum_t x_i(t)^2 - \left[\frac{1}{5000} \sum_t x_i(t) \right]^2 
\end{eqnarray*}

@@

Due to computational restraints, which we discuss further in the
following section, we are only able to increase $p$ to a maximum value
of 6 and must also restrict to $N = 2$. We maintain $\alpha = 0.02$
and $\tau = 0.05$ to match with the parameters used in Figure
\ref{fig:pvariation}. Finally, in Figure \ref{fig:pVarianceExpt}, we
are averaging over 10 choices of $\Gamma$ in the range $[-1, 0]$.
